Using Genetic Algorithms to Optimize the Analog Method for Precipitation Downscaling in the Swiss Alps

### 2016-12-01 Optimization case study - JoH/elsarticle-template.tex
JoH paper under correction.

\begin{abstract}
The Analog Method aims at predicting precipitation based on predictors variables provided by global models. The method samples a selection of days in the archives that are similar to the target day to predict, and consider the set of their corresponding observed precipitation (the predictand) as the conditional distribution for the target day. The relationship between predictors and predictands relies on some parameters that characterize how and where the similitude between two atmospheric situations is defined.

This relationship is usually established by a semi-automatic sequential procedure that has strong limitations. A new global optimization approach relying on Genetic Algorithms can optimize all parameters jointly and automatically, which is a breakthrough in the way the Analog Method was calibrated until now. It allows taking into account parameters inter-dependencies, and selecting objectively some parameters that were manually selected beforehand (such as the pressure levels and the temporal windows of the predictor variables).

@@

The analog method (AM) is a downscaling technique based on the idea expressed by \citet{Lorenz1956, Lorenz1969}, that similar situations in terms of atmospheric circulation are likely to lead to similar local weather \citep{Duband1970, Bontron2005}. It aims at predicting a predictand, often the daily precipitation \citep[eg.][]{Guilbaud1997, Bontron2005, Bliefernicht2010, Marty2012, Horton2012, Radanovics2013, BenDaoud2015}, on the basis of predictor variables describing the synoptic atmospheric circulation. Other predictands are also often considered \citep[see][for a non-exhaustive list]{Horton2016}.

In operational forecasting, it has been used mainly by practitioners, notably hydropower companies or flood forecasting services, while in climate studies, it is used to downscale the results of global climate model (GCM) simulation runs. In this last case, although the GCM represents the large or regional scale evolutions of the atmosphere, it would be far too computer intensive to get down to variables such as precipitation at a rather small and representative scale. Therefore the idea is to rely on past observations to bypass these unaffordable small scale simulations and to go from the large scale situation proposed by the GCM to the end variables like precipitation by searching analogs in the archives.

The method can be made of multiple successive subsampling steps, each of them relying on different meteorological variables. A certain number of parameters define the relationship between predictors and predictands, such as the choice of the predictor variable, its pressure level and temporal window to consider, the spatial domain to use for the comparison, as well as the analogy criteria itself, and finally the number of analog situations to keep at each subsampling level. These parameters are usually calibrated by means of a semi-automatic sequential procedure \citep[see][for the details]{Bontron2004, Horton2016}, i.e. by optimizing each single parameter one at a time, in an arbitrarily chosen order, with no or little reconsidering. This sequential approach has therefore strong limitations: (i) it cannot automatically choose the pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters within a level of analogy, and even less between them, and (iii) it could not easily handle new degrees of freedom, such as a weighting between the pressure levels. Thus, even if the processing involved is relatively fast, the sequential approach requires laborious assessments of predictors combinations (variables, pressure levels, temporal windows), and present a high risk of ending in a local optimum due to subjective initial choices and lack of consideration of parameters inter-dependencies.

With the perspective to overcome these limitations, a global optimization by Genetic Algorithms (GAs) was introduced by \citet{Horton2016a}. An intensive assessment work resulted in recommendations of parametrization of GAs in order to optimize AMs successfully. The present paper is based on these recommendations and illustrate them on precipitation predicting for the upper Rh\^{o}ne catchment in the Swiss Alps. It aims at proving the relevance of a fully-automatic, objective, and global, optimization technique for AMs. The applications are indeed numerous, as the AM has to be adapted to every new location it is applied, or to any new predictand it should predict.

A short overview of AMs is presented in section \ref{sec:analog_method}, as well as a summary of Genetic Algorithms in section \ref{sec:gas}. Section \ref{sec:case_study} describes the case study area. The results are first detailed for the optimization of the analogy on the atmospheric circulation only (section \ref{sec:optim_circul}), before being extended to a method adding a second level of analogy on moisture variables (section \ref{sec:optim_moisture}). General discussions (section \ref{sec:discussion}) and conclusions (section \ref{sec:conclusions}) follow.

@@

Multiple variations of the methods exist, and most of them will not be detailed hereafter \cite[see][for more comprehensive listings]{Horton2016, BenDaoud2015}. However, there are mainly 2 parameterizations that are most often used for precipitation downscaling and that will be considered as reference: one that relies on an analogy of the atmospheric circulation, and another that adds a second level of analogy on moisture variables \citep{Obled2002, Bontron2005, Marty2012}.

The method based on the analogy of the synoptic circulation consists in the following steps (Table \ref{table:params_R1}): the similarity of the atmospheric circulation of a target date with every day of the archive is assessed by processing the S1 criteria \citep[Eq.\ \ref{eq:S1}, ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a certain spatial window. \citet{Bontron2005} showed that the geopotential height at 500~hPa (Z500) and 1000~hPa (Z1000) are the best first predictors of the NCEP/NCAR reanalysis dataset, and that the S1 criteria performs better than scores based on absolute distances. The reason for such better results is that the S1 criteria allows comparing the circulation pattern, by means of the gradients, rather than the absolute value of the geopotential height. To cope with seasonal effects, candidate dates are extracted within a period of 4 months centered around the target date, for every year of the archive. Following the nomenclature proposed by \citet{Horton2016}, this method using two geopotential heights will be named 2Z.

\begin{equation}
\label{eq:S1}
S1=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
{\displaystyle \sum_{i} max\left\lbrace \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert \right\rbrace }
\end{equation}
where $\Delta \hat{z}_{i}$ is the difference in geopotential height between the \textit{i}th pair of adjacent points of the gridded data describing the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both directions over the selected spatial domain. The smaller the S1 values, the more similar the pressure fields.

@@

The other most know parametrization adds a second level of analogy on moisture variables (method 2Z-2MI, Table \ref{table:params_R2}). The predictor that \citet{Bontron2004} found optimal for the France territory is a moisture index made of the product of the precipitable water with the relative humidity at 850~hPa (RH850). \cite{Horton2012a} confirmed that this index is also better for the Swiss Alps than any other variable from the NCEP/NCAR reanalysis considered independently. When adding a second level of analogy, $N_{2}$ dates are subsampled in the $N_{1}$ analogs on the atmospheric circulation, to end up with a smaller number of analog situations. When a second level of analogy is added, a higher number of analogs $N_{1}$ is kept on the first level.

@@

Predictors are generally reanalysis datasets. NCEP-NCAR reanalysis \citep[6-hourly, 17 pressure levels at a resolution of 2.5\degree, see][]{Kalnay1996} are used here, but it could be any other reanalysis dataset.

The predictand (which is to be predicted) is here the daily precipitation (6~a.m. to 6~a.m. the next day) measured at the MeteoSwiss' stations network, for the period 1961-2008. The time series from every available gauging station were averaged over subregions of approximately 500~km$^{2}$ in order to smooth local effects \citep{Obled2002, Marty2012}.

@@

The score that is most often used to assess an AM performance is the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analog situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. The mean CRPS of a prediction series of length $n$ can be written:

@@

In order to compare the value of the score in regard to a reference, one often considers its skill score expression, and use the climatological distribution of daily precipitation as the reference. The CRPSS (\textit{Continuous Ranked Probability Skill Score}) is thus defined as following:

@@

Genetic Algorithms (GAs) were developed by \citet{Holland1992b} and \citet{Goldberg1989}. They are part of Evolutionary Algorithms \citep{Back1993b, Schwefel1993}, which get inspiration from some mechanisms of biological evolution, such as reproduction, genetic mutations, chromosomal crossovers, and natural selection. GAs seek the global optimum on a complex surface, theoretically without restriction. Practically, GAs allow rapidly approaching satisfactory solutions, but they do not provide the optimum solution for sure \citep{Zitzler2004a}. It is indeed mainly a matter of time. When the optimizer gets closer to the global optimum, any new improvement takes more time to appear (see for example Figure \ref{fig:evolution}), and the final adjustment of the parameters is very time consuming \citep{Back1993a}. For problems that require a significant amount of time in order to evaluate the objective function, as in the case of AMs, the number of generations has to be limited in order to get reasonable processing time. Thus, different acceptable solutions can result from one or more optimizations \citep{Holland1992b}. This is a strength and a weakness of GAs: they are very good at exploring complex parameter spaces in order to identify the most promising areas, but they will not necessarily always find the best solution with the optimal values of all parameters \citep{Holland1992b}.

The optimizations were performed based on the recommended GAs parametrization for AMs as described in \citet{Horton2016a}. As the optimization is mostly sensitive to the mutation operator (that randomly changes some values in the parameters sets), parallel optimizations are considered with variants of this operator, according to \citet{Horton2016a}:

\begin{itemize}
	\item 3x non-uniform mutation \citep{Michalewicz1996} with varying parameters,
	\item 1x multi-scale mutation \citep{Horton2016a},
	\item 2x chromosome of adaptive search radius \citep{Horton2016a}
\end{itemize}

@@

The study area is the alpine upper Rh\^{o}ne catchment in Switzerland (Fig.\ \ref{fig:map}). The altitude ranges from 372 to 4634~m.a.s.l.\ and the area is 5524~km$^{2}$. This region is the target of the MINERVE (Mod\'{e}lisation des Intemp\'{e}ries de Nature Extr\^{e}me sur les Rivi\`{e}res Valaisannes et de leurs Effets) project that aims at allowing real-time flood management on the upper Rh\^{o}ne catchment \citep{GarciaHernandez2009b}. Even though the region is rather small, the meteorological influences related to extreme weather conditions varies substantially within it \citep[see][]{Horton2012}. Based on different climatological analyses, the gauging stations in the catchment were clustered in 10~subregions (Fig.\ \ref{fig:map}) :

\begin{enumerate}
	\item Swiss Chablais
	\item Trient Valley
	\item West Bernese Alps
	\item Lower Rhone Valley
	\item Left side valleys
	\item Southern ridges
	\item Upper Rhone Valley
	\item Southeast ridges
	\item East Bernese Alps
	\item Conches Valley
\end{enumerate}

@@

The analogy of the atmospheric circulation was optimized for the 10 subregions  independently. We started from the most simple AM, and increased the complexity in order to identify the degrees of freedom that are of particular interest. Thus, the tested parametrization evolved iteratively in complexity. The detailed results of the intermediate stages are not provided \citep[see][for the details]{Horton2012a}.

The reference method for the analogy of the atmospheric circulation (2Z, Table \ref{table:params_R1}), based on Z500 and Z1000, was first considered. The optimizer had to choose the number of analogs, both spatial windows with no overlapping constraint, as well as the temporal windows (hours of observation of the geopotential), what the sequential calibration cannot do. With these degrees of freedom, a relative CRPSS improvement of 3.97\% and 2.45\% in average was obtained for the CP and the VP respectively. Some tests showed that most of the gains are due to the non-overlapping spatial windows. This is not a tremendous improvement, but it demonstrated that the optimizer was able to get relevant parameters.

Then, an additional degree of freedom was provided to the GAs by letting them choose the pressure levels along with the other parameters, which is also a non-automated process in the sequential calibration. This degree of freedom increased the optimization time and may decrease the number of simulations that converge to a single solution. However, most solutions were very close in terms of score. The averaged relative improvement of the CRPSS is 5.63\% for the CP and 3.82\% for the VP. The pressure levels that were chosen are Z500 or Z700 for the upper level, and Z925 or Z1000 (most often) for the lower level.

Parallel analyses showed that the analogy of circulation is incomplete, and that the geopotential still contains relevant information that can improve the statistical relationship. Therefore, a third predictor was added (still on the geopotential height) that the optimizer could use along with the previous parameters. There was no constraint on the predictors, so that the same pressure level could be selected more than once. Some improvements were found on the score, both for the CP and the VP, confirming that this additional information is beneficial for the quality of the prediction. We then tried with 4 predictors, and so on, up to 8. Every time a new predictor to optimize was added, the score on the CP increased, but always more to a smaller extent. However, the score value on the VP dropped after 4 predictors, revealing an over-parametrization of the method, and thus a lack of robustness. Considering 4 predictors is optimal for this case study, since the gain in CRPSS is significant. It cannot be excluded that another number would prevail in another region under other meteorological conditions.

Finally, a weighting of the criteria values per pressure level was proposed, again optimized by GAs. The weighting operates in the combination of the S1 criteria processed on every level, which were previously averaged with equal weights. The role of this new degree of freedom is to give more weight to the levels with greater predictive capacity, and to consider the geopotential variability changes with altitude.

@@

The chosen method for the atmospheric circulation analogy, based on 4 levels of the geopotential, and that will be name 4Zo, is made of the following degrees of freedom:

@@

To assess the parameters cross-compatibility, those optimized for one subregion were applied to the others. The resulting losses or gains of the CRPSS are displayed in Figure \ref{fig:crossing_4Zo_calib} for the CP and in Figure \ref{fig:crossing_4Zo_valid} for the VP.

@@

The automatic selections of the pressure levels (Table \ref{table:levels_GA_4Zo}) and the temporal windows (not shown) for the analogy of circulation show a great homogeneity and are spatially consistent. First of all, the level Z1000 is always selected twice (the first time at 6 or 12~h, and the second always at 30~h) and Z700 is selected once for every subregion (always at 24~h). The level which varies from one subregion to another is the upper level (however always at 12~h), which is Z300 for the North-West part of the catchment, Z500 for most of the subregions, and Z600 for the Conches Valley. Its spatial distribution is however homogeneous. The optimizer thus provided consistent selections of pressure levels and temporal windows, which depicts a significant preference in the AM, and the success of GAs to provide consistent results. The automatic selection of the pressure levels is a big advantage in favor of a global optimization.

The resulting spatial windows (Figure \ref{fig:spatial_windows_4Zo}) may look very diverse first, but there are significant similarities for subregions located within the same vicinity. The first 4 subregions are characterized by a large spatial window on the upper level, whereas it is smaller elsewhere. For most subregions, the second level (Z700) is represented by thin and longitudinally extended domains. The third level (Z1000 at 6 or 12~h) also contains longitudinally extended domains, but a bit larger. The last one (Z1000 at 30~h) is rather large and squared. Subregions number 5 (Left side valleys) and 6 (Southern ridges) have exactly the same spatial windows, which suggest that they behave in a similar way and thus could have been merged. This similarity is a good sign for the accuracy of the optimized parameters.

@@

When optimizing a method made of 2 levels of analogy, the introduction of moisture variables in the second level has an influence on the parameter values of the first level. This is first visible on the number $N_{1}$ of analogs to be selected on the first level, and on the selection of the pressure levels for the circulation analogy.

@@

As for the pressure levels, Z1000 that was previously systematically selected twice (Table \ref{table:levels_GA_4Zo}) is now less chosen (once or even not at all) for both 4Zo-2MIo (Table \ref{table:levels_GA_4Zo_2MIo}) and 4Zo-4MIo (not shown). There is indeed a shift of the previously selected Z1000 for higher levels, that is even slightly stronger with 4 moisture predictors than with 2. This change is likely due to the fact that when considering only the circulation analogy, the method tries to take into account information that can serve as proxy for moisture assessment, whereas it does not need it with the moisture index. This aspect has never been demonstrated before, as sequential calibration tools do not allow it. It can only be assessed by a global optimization technique that can tune jointly both levels of analogy.

@@

The convergence of parallel optimizations decreases when the method to optimize becomes more and more complex. The optimizer do not always converge to the exact global optimum, but to its surroundings. This is related to the fact that the optimization slows down when it gets closer to the global optimum, and that one has to stop it before the end, due to the required processing time (see Figure \ref{fig:evolution} as example). The resulting parameters may sometimes present non negligible differences, even though the score is almost similar. Through some Monte-Carlo analyses of the parameter space properties of the AM, \citet{Horton2012a} showed that some parameters of the method have a wide range of acceptable values \citep[see also][]{Horton2016}. The spatial windows, for example, can be larger than the optimal size without much impact on the score, while they cannot be smaller \citep[see also][]{Bontron2004}. We could also observe that the selection of the pressure level is not a parameter as discrete as we would have thought, and that choosing another level may have reduced consequences on the performance. This is particularly true for higher pressure levels but can be more critical for lower layers. It was thus interesting to sometimes get several sets of near-optimal parameters, but with some nuances, in order to get an idea of the sensitivity of the parameters for a given region, and to compare the score on the VP. In this regard, a cross-validation technique may be advisable. An approach that can also be recommended is to first explore a wide range of the parameter space with some optimizations, and to narrow it according to the results for more targeted optimizations that are likely to go faster and to perform better.

@@

GAs are relatively heavy on processing and require an IT infrastructure capable of performing thousands hours of calculations. However, they automatically optimize all parameters of the AM, what the sequential calibration does not allow. Therefore, much human time is saved, that was previously required to assess successively numerous combinations of parameters (particularly the selection of the pressure levels and the temporal windows). The ability to optimize jointly all parameters is important given the strong dependencies between them and between the levels of analogy.

>>>>

\begin{abstract}
Analog methods aim at predicting precipitation based on synoptic predictors provided by general circulation models or numerical weather prediction models. The method samples a selection of days in the archives that are similar to the target day to predict, and consider the set of their corresponding observed precipitation (the predictand) as the conditional distribution for the target day. The relationship between predictors and predictands relies on some parameters that characterize how and where the similitude between two atmos+pheric situations is defined.

This relationship is usually established by a semi-automatic sequential procedure that has strong limitations: (i) it cannot automatically choose the pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters within a level of analogy, and even less between them, and (iii) it could not easily handle new degrees of freedom. A new global optimization approach relying on genetic algorithms can optimize all parameters jointly and automatically, which is a breakthrough in the way the analog method was calibrated until now. It allows taking into account parameters inter-dependencies, and selecting objectively some parameters that were manually selected beforehand (such as the pressure levels and the temporal windows of the predictor variables).

@@

The analog method (AM) is a downscaling technique based on the idea expressed by \citet{Lorenz1956, Lorenz1969}, that similar situations in terms of atmospheric circulation are likely to lead to similar local weather \citep{Duband1970}. It uses predictor variables describing the synoptic atmospheric circulation in order to predict local-scale predictands of interest. It is often used to predict daily precipitation, either in an operational forecasting context \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Hamill2015, BenDaoud2016}, or a climate context \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015, Raynaud2016b}. Other predictands are also considered, such as precipitation radar images \citep{Panziera2011,Foresti2015a}, temperature \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016, Raynaud2016b}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, and radiation \citep{Bois1981, Raynaud2016b}.

In operational forecasting, it has been used mainly by practitioners, notably hydropower companies or flood forecasting services, while in climate studies, it is used to downscale the results of global climate model (GCM) simulation runs. In this last case, although the GCM represents the large or regional scale evolutions of the atmosphere, it would be far too computer intensive to resolve variables such as precipitation at a rather small and representative scale. Therefore the idea is to rely on past observations to bypass these unaffordable small scale simulations and to go from the large scale situation proposed by the GCM to the end variables like precipitation by searching analogs in the archives.

The method can be designed with multiple successive subsampling steps, each of them relying on different meteorological variables. A certain number of parameters define the relationship between predictors and predictands, such as the choice of the predictor variable, its pressure level and temporal window to consider, the spatial domain to use for the comparison, as well as the analogy criterion itself, and finally the number of analog situations to keep at each subsampling level. These parameters are usually calibrated by means of a semi-automatic sequential procedure \citep[see][for the details]{Bontron2004, Radanovics2013, Horton2016a}, i.e. by optimizing each single parameter one at a time, in an arbitrarily chosen order, with no or little reassessment. This sequential approach has therefore strong limitations: (i) it cannot automatically choose the optimal pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters within a level of analogy, and even less between them, and (iii) it could not easily handle new degrees of freedom, such as a weighting between the pressure levels. Thus, even if the processing involved is relatively fast, the sequential approach requires laborious assessments of predictors combinations (variables, pressure levels, temporal windows), and presents a high risk of ending in a local optimum due to subjective initial choices and lack of consideration of parameter inter-dependencies. Other calibration methods exist for specific applications, such as the radar images \citep{Panziera2011, Foresti2015a}.

With the perspective to overcome these limitations, a global optimization by genetic algorithms (GAs) was introduced by \citet{Horton2016}. An intensive assessment work resulted in recommendations of parametrization of GAs in order to optimize AMs successfully. The present paper is based on these recommendations and illustrate them for precipitation prediction for the upper Rh\^{o}ne catchment in the Swiss Alps. It aims at proving the relevance of a fully-automatic, objective, and global, optimization technique for AMs. The applications are indeed numerous, as the AM has to be adapted to every new location it is applied, or to any new predictand it should predict.

A short overview of AMs is presented in section \ref{sec:analog_method}, as well as a summary of genetic algorithms in section \ref{sec:gas}. Section \ref{sec:case_study} describes the case study area. The results are first detailed for the optimization of the analogy on the atmospheric circulation only (section \ref{sec:optim_circul}), before being extended to a method adding a second level of analogy on moisture variables (section \ref{sec:optim_moisture}). General discussions (section \ref{sec:discussion}) and conclusions (section \ref{sec:conclusions}) follow.

@@

Multiple variations of the methods exist, and most of them will not be detailed hereafter \cite[see][for more comprehensive listings]{BenDaoud2016}. However, there are mainly 2 parameterizations that are most often used for precipitation downscaling and that will be considered as reference: one that relies on an analogy of the atmospheric circulation, and another that adds a second level of analogy on moisture variables \citep{Obled2002, Bontron2005, Marty2012}.

The method based on the analogy of the synoptic circulation consists in the following steps (Table \ref{table:params_R1}): the similarity of the atmospheric circulation of a target date with every day of the archive is assessed by processing the S1 criterion \citep[Eq.\ \ref{eq:S1}, ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a certain spatial window:

\begin{equation}
\label{eq:S1}
S1=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
{\displaystyle \sum_{i} max\left\lbrace \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert \right\rbrace }
\end{equation}
where $\Delta \hat{z}_{i}$ is the difference in geopotential height between the \textit{i}th pair of adjacent points of the gridded data describing the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both North and East directions over the selected spatial domain. The smaller the S1 values, the more similar the pressure fields.

\citet{Bontron2005} showed that the geopotential height at 500~hPa (Z500) and 1000~hPa (Z1000) are the best first predictors of the NCEP/NCAR reanalysis dataset, and that the S1 criterion performs better than scores based on absolute distances. The reason for such better results is that the S1 criterion allows comparing the circulation pattern, by means of the gradients, rather than the absolute value of the geopotential height, which represent better the flow direction. To cope with seasonal effects, candidate dates are extracted within a period of 4 months centered around the target date, for every year of the archive. This method using two geopotential heights will be named 2Z.

@@

The other most know parametrization adds a second level of analogy on moisture variables (method 2Z-2MI, Table \ref{table:params_R2}). The predictor that \citet{Bontron2004} found optimal for the France territory is a moisture index made of the product of the precipitable water with the relative humidity at 850~hPa (RH850). \cite{Horton2012a} confirmed that this index is also better for the Swiss Alps than any other variable from the NCEP/NCAR reanalysis considered independently. When adding a second level of analogy, $N_{2}$ dates are subsampled in the $N_{1}$ analogs on the atmospheric circulation, to end up with a smaller number of analog situations. When a second level of analogy is added, a higher number of analogs $N_{1}$ is kept on the first level. Moisture fields are not as well predicted by NWP models as pressure variables. It means that the 2Z-2MI method, when used in real-time forecasting, is very dependent on the skill of the NWP model in predicting moisture fields, and thus is often used on limited lead times.

@@

Predictors are generally extracted from reanalysis datasets. NCEP-NCAR reanalysis \citep[6-hourly, 17 pressure levels at a resolution of 2.5\degree, see][]{Kalnay1996} are used here, but it could be any other reanalysis dataset.

The predictand (which is to be predicted) is here the daily precipitation (6~a.m. to 6~a.m. the next day) measured at the MeteoSwiss' stations network, for the period 1961-2008. The time series from every available gauging station were averaged over subregions of approximately 500~km$^{2}$ in order to smooth local effects \citep{Obled2002, Marty2012}. It helps accounting for local variability, mainly when convective processes are involved, which increases slightly the prediction skill.

@@

The performance assessment in the present context consists in verifying the prediction of an ensemble-probabilistic technique. The score that is most often used to assess an AM performance is the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analog situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. The mean CRPS of a prediction series of length $n$ can be written:

@@

In order to compare the value of the score in regard to a reference, one often considers its skill score expression, and use the climatological distribution of precipitation from the entire archive as the reference. The CRPSS (\textit{Continuous Ranked Probability Skill Score}) is thus defined as following:

@@

Genetic algorithms (GAs) were developed by \citet{Holland1992b} and \citet{Goldberg1989}. They are part of Evolutionary Algorithms \citep{Back1993b, Schwefel1993}, which get inspiration from some mechanisms of biological evolution, such as reproduction, genetic mutations, chromosomal crossovers, and natural selection. GAs seek the global optimum on a complex surface, theoretically without restriction, which is of interest for AMs, that are characterized by a complex high-dimensional error function having multiple local optima. Practically, GAs allow rapidly approaching satisfactory solutions, but they do not provide the optimum solution for sure \citep{Zitzler2004a}. It is indeed mainly a matter of time. When the optimizer gets closer to the global optimum, any new improvement takes more time to appear (see for example the slow-down of the improvements over generations in Figure \ref{fig:evolution}), and the final adjustment of the parameters is very time consuming \citep{Back1993a}. For problems that require a significant amount of time in order to evaluate the objective function, as in the case of AMs, the number of generations has to be limited in order to get reasonable processing time. Thus, different acceptable solutions can result from one or more optimizations \citep{Holland1992b}. This is a strength and a weakness of GAs: they are very good at exploring complex parameter spaces in order to identify the most promising areas, but they will not necessarily always find the best solution with the optimal values of all parameters \citep{Holland1992b}.

The optimizations were performed based on the recommended GAs parametrization for AMs as described in \citet{Horton2016}. As the optimization is mostly sensitive to the mutation operator (that randomly changes some values in the parameters sets), parallel optimizations are considered with variants of this operator, according to \citet{Horton2016}:

\begin{itemize}
	\item 3x non-uniform mutation \citep{Michalewicz1996} with varying parameters,
	\item 1x multi-scale mutation \citep{Horton2016},
	\item 2x chromosome of adaptive search radius \citep{Horton2016}
\end{itemize}

@@

The study area is the alpine upper Rh\^{o}ne catchment in Switzerland (Fig.\ \ref{fig:map}). The altitude ranges from 372 to 4634~m.a.s.l.\ and the area is 5524~km$^{2}$. This region is the target of the MINERVE (Mod\'{e}lisation des Intemp\'{e}ries de Nature Extr\^{e}me sur les Rivi\`{e}res Valaisannes et de leurs Effets) project that aims at allowing real-time flood management on the upper Rh\^{o}ne catchment \citep{GarciaHernandez2009b}. Even though the region is rather small, the meteorological influences related to extreme weather conditions varies substantially within it \citep[see][]{Horton2012}. Indeed, a high spatial variability exists, which is due to the complex orography of the region. Based on different climatological analyses, the gauging stations in the catchment were clustered in 10~subregions (Fig.\ \ref{fig:map}) :

\begin{enumerate}
	\item Swiss Chablais
	\item Trient Valley
	\item West Bernese Alps
	\item Lower Rhone Valley
	\item Southern valleys
	\item Southern ridges
	\item Upper Rhone Valley
	\item Southeast ridges
	\item East Bernese Alps
	\item Conches Valley
\end{enumerate}

@@

The analogy of the atmospheric circulation was optimized for the 10 subregions  independently. We started from the most simple AM, and increased the complexity in order to identify the degrees of freedom that are of particular interest. Thus, the tested parametrization evolved iteratively in complexity. The detailed results of the intermediate stages are not provided in this paper \citep[see][for the details]{Horton2012a}.

The reference method for the analogy of the atmospheric circulation (2Z, Table \ref{table:params_R1}), based on Z500 and Z1000, was first considered. The optimizer had to choose the number of analogs, both spatial windows with no overlapping constraint, as well as the temporal windows (hours of observation of the geopotential), what the sequential calibration cannot do. With these degrees of freedom, a relative CRPSS improvement of 3.97\% and 2.45\% in average was obtained for the CP and the VP respectively, with regard to the reference method based on Z500 and Z1000. Some tests showed that most of the gains are due to the non-overlapping spatial windows. This is not a large improvement, but it demonstrated that the optimizer was able to get relevant parameters.

Then, an additional degree of freedom was provided to the GAs by letting them choose the pressure levels along with the other parameters (analog numbers, spatial and temporal windows), which is also a non-automated process in the sequential calibration. This degree of freedom increased the optimization time and may decrease the number of simulations that converge to a single solution. However, most solutions were very close in terms of score. The averaged relative improvement of the CRPSS is 5.63\% for the CP and 3.82\% for the VP. The pressure levels that were chosen are Z500 or Z700 for the upper level, and Z925 or Z1000 (most often) for the lower level.

Parallel analyses showed that the analogy of circulation is incomplete, and that the geopotential still contains relevant information that can improve the statistical relationship. Therefore, a third predictor was added (still on the geopotential height) that the optimizer could use along with the previous parameters. There was no constraint on the predictors, so that the same pressure level could be selected more than once. Some improvements were found on the score, both for the CP and the VP, confirming that this additional information is beneficial for the quality of the prediction. We then tried with 4 predictors, and so on, up to 8. Every time a new predictor to optimize was added, the score on the CP increased, but in a reduced amount. However, the score value on the VP dropped after 4 predictors, revealing an over-parametrization of the method, and thus a lack of robustness. Considering 4 predictors is optimal for this case study, since the gain in CRPSS is significant. It cannot be excluded that another number would prevail in another region than the upper Rh\^{o}ne catchemnt, under other meteorological conditions.

Finally, a weighting of the analogy criteria values per pressure level was proposed, again optimized by GAs. The weighting operates in the combination of the S1 criteria processed on every level, which were previously averaged with equal weights. The role of this new degree of freedom is to give more weight to the levels with greater predictive capacity, and to consider the geopotential variability changes with altitude.

@@

The chosen method for the atmospheric circulation analogy, based on 4 levels of the geopotential, and that will be named 4Zo, is made of the following degrees of freedom:

@@

To assess the parameters cross-compatibility, and the spatial coherence of the resulting parameters, those optimized for one subregion were applied to the others. The resulting losses or gains of the CRPSS are displayed in Figure \ref{fig:crossing_4Zo_calib} for the CP and in Figure \ref{fig:crossing_4Zo_valid} for the VP.

@@

The automatic selections of the pressure levels (Table \ref{table:levels_GA_4Zo}) and the temporal windows (not shown) for the analogy of circulation show a great homogeneity and are spatially consistent. First of all, the level Z1000 is always selected twice (the first time at 6 or 12~h, and the second always at 30~h) and Z700 is selected once for every subregion (always at 24~h). The level which varies from one subregion to another is the upper level (however always at 12~h), which is Z300 for the North-West part of the catchment, Z500 for most of the subregions, and Z600 for the Conches Valley. Its spatial distribution is however homogeneous. The optimizer thus provided consistent selections of pressure levels and temporal windows. The automatic selection of the pressure levels is a big advantage in favor of a global optimization.

The resulting spatial windows (Figure \ref{fig:spatial_windows_4Zo}) may look very diverse first, but there are significant similarities for subregions located within the same vicinity. The first 4 subregions are characterized by a large spatial window on the upper level, whereas it is smaller elsewhere. For most subregions, the second level (Z700) is represented by thin and longitudinally extended domains. The third level (Z1000 at 6 or 12~h) also contains longitudinally extended domains, but a bit larger. The last one (Z1000 at 30~h) is rather large and squared. Subregions number 5 (Southern valleys) and 6 (Southern ridges) have exactly the same spatial windows, which suggest that they behave in a similar way and thus could have been merged. This similarity is a good sign for the accuracy of the optimized parameters.

@@

When optimizing a method made of 2 levels of analogy, the introduction of moisture variables in the second level has an influence on the parameter values of the first level. It means that the two levels of analogy bring complementary information, and are thus not independent. This is first visible on the number $N_{1}$ of analogs to be selected on the first level, and on the selection of the pressure levels for the circulation analogy.

@@

As for the pressure levels, Z1000 that was previously systematically selected twice (Table \ref{table:levels_GA_4Zo}) is now less chosen (once or even not at all) for both 4Zo-2MIo (Table \ref{table:levels_GA_4Zo_2MIo}) and 4Zo-4MIo (not shown). There is indeed a vertical shift of the previously selected Z1000 for higher levels, that is even slightly stronger with 4 moisture predictors than with 2. This change is likely due to the fact that when considering only the circulation analogy, the method tries to take into account information that can serve as proxy for moisture assessment, whereas it does not need it with the moisture index. This aspect has never been demonstrated before, as sequential calibration tools do not allow it. It can only be assessed by a global optimization technique that can tune jointly both levels of analogy.

@@

The convergence of parallel optimizations decreases when the method to optimize becomes more and more complex. The optimizer does not always converge to the exact global optimum, but to its surroundings. This is related to the fact that the optimization slows down when it gets closer to the global optimum, and that one has to stop it before the end, due to the required processing time (see Figure \ref{fig:evolution} as example). The resulting parameters may sometimes present non negligible differences, even though the score is almost similar. Through some Monte-Carlo analyses of the parameter space properties of the AM, \citet{Horton2012a} showed that some parameters of the method have a wide range of acceptable values. The spatial windows, for example, can be larger than the optimal size without much impact on the score, while they cannot be smaller \citep[see also][]{Bontron2004}. We could also observe that the selection of the pressure level is not a parameter as discrete as we would have thought, and that choosing another level may have reduced consequences on the performance. This is particularly true for higher pressure levels but can be more critical for lower layers. It was thus interesting to sometimes get several sets of near-optimal parameters, but with some nuances, in order to get an idea of the sensitivity of the parameters for a given region, and to compare the score on the VP. In this regard, a cross-validation technique may be advisable. An approach that can also be recommended is to first explore a wide range of the parameter space with some optimizations, and to narrow it according to the results for more targeted optimizations that are likely to go faster and to perform better.

@@

GAs are relatively heavy on processing and require an IT infrastructure capable of performing thousands hours of calculations. However, they automatically optimize all parameters of the AM, what is not possible with the sequential calibration. Therefore, much human time is saved, that was previously required to assess successively numerous combinations of parameters (particularly the selection of the pressure levels and the temporal windows). The ability to optimize jointly all parameters is important given the strong dependencies between them and between the levels of analogy.
